Is high social class always beneficial for survival? Northern Sweden 1801--2013

### 2017-11-28 article2.Rmd
manipulating!

---
title: Life course and long-term perspectives of social inequality in mortality among
  elderly and adults in Northern Sweden      1801–2013
author: "S&ouml;ren Edvinsson and G&ouml;ran Brostr&ouml;m"
date: '`r Sys.Date()`'
output:
  bookdown::pdf_document2:
    citation_package: natbib
    keep_tex: yes
    number_sections: yes
    toc: yes
    toc_depth: 2
  bookdown::html_document2:
    number_sections: yes
    toc: yes
    toc_depth: 2
    toc_float: yes
  bookdown::word_document2:
    toc: yes
    toc_depth: 2
bibliography: mybib.bib
documentclass: article
fontsize: 12pt
citation_package: natbib
biblio-style: apalike
---

>>>>

---
title: "Life course and long-term perspectives of social inequality in mortality among
  elderly and adults in Northern Sweden 1801–2013"
author: "S&ouml;ren Edvinsson and G&ouml;ran Brostr&ouml;m"
date: '`r Sys.Date()`'
output:
  bookdown::pdf_document2:
    citation_package: natbib
    keep_tex: yes
    number_sections: yes
    toc: yes
    toc_depth: 2
  bookdown::html_document2:
    number_sections: yes
    toc: yes
    toc_depth: 2
    toc_float: yes
  bookdown::word_document2:
    toc: yes
    toc_depth: 2
bibliography: mybib.bib
documentclass: article
fontsize: 12pt
citation_package: natbib
biblio-style: apalike
---

### 2017-11-28 article3.Rmd
manipulating!

```{r readalldata}
source("R/get.file.R")
skel <- get.file("skel")
##
skume <- get.file("skume")
##
foli <- get.file("foli")
skel$urban <- FALSE
vb0Tab <- rbind(skel, skume[, names(skel)], foli[, names(skel)])
vb0Tab$civst <- relevel(vb0Tab$civst, ref = "married")
```

@@

```{r getske4060}
cap1w <-  "Cumulative hazards for social class by time period, ages 40--64, women."
scap1w <-  "Cumulative hazards, ages 40--64, women."
cap1m <-  "Cumulative hazards for social class by time period, ages 40--64, men."
scap1m <-  "Cumulative hazards, ages 40--64, men."
##
cap0 <- "Probability of dying before age 65 for a 40-year-old by social class and period, women (top) and men (bottom)."
scap0 <- "Probability of dying before age 65 for a 40-year old by social class."
cap12 <- "Relative probability of dying before age 65 for a 40-year-old by social class and period, women (top) and men (bottom)."
scap12 <- "Relative probability of dying before age 65 for a 40-year-old by social class."
cap12log <- "Probability of dying before age 65 for a 40-year-old by social class and period, women (top) and men (bottom). The last decades."
scap12log <- "Probability by social class, ages 40--64, last years."

@@

```{r hazcalw}
source("R/getPHglm2.R")
##nper <- 4
nper <- length(labb)
resw <- getPHglm2(vb4064[vb4064$sex == "female", ],
            nper, labb)
resm <- getPHglm2(vb4064[vb4064$sex == "male", ],
            nper, labb)
```

@@

```{r cumswstand, fig.cap = cap12, fig.scap = scap12, fig.height = 8, out.height = "14cm"}
oldpar <- par(mfrow = c(2, 1))
woutp <- plotED.glm(resw, col = scol, logScale = FALSE, relative = TRUE,
                    ylab = "P(dead at 65 | alive at 40)", ylim = c(0.1, 0.55), main = "Women 40-64")##, exclude = 5:6)
moutp <- plotED.glm(resm, col = scol, logScale = FALSE, relative = TRUE,
                    ylab = "P(dead at 65 | alive at 40)", ylim = c(0.1, 0.55), main = "Men 40-64")##, exclude = 5:6)
par(oldpar)
```

@@

```{r cumswold, fig.cap = cap22, fig.scap = scap22, fig.height = 8, out.height = "14cm"}
##source("R/plotED.R")
resw <- getPHglm2(vb6589[vb6589$sex == "female", ],
            nper, labb)
resm <- getPHglm2(vb6589[vb6589$sex == "male", ],
            nper, labb)

oldpar <- par(mfrow = c(2, 1))
woutp <- plotED.glm(resw, col = scol, logScale = FALSE, relative = TRUE,
                    ylab = "P(dead at 90 | alive at 65)", ylim = c(0.1, 0.55), main = "Women 65-89")##, exclude = 5:6)
moutp <- plotED.glm(resm, col = scol, logScale = FALSE, relative = TRUE,
                    ylab = "P(dead at 90 | alive at 65)", ylim = c(0.1, 0.55), main = "Men 65-89")##, exclude = 5:6)
par(oldpar)
```

>>>>

```{r readalldata}
source("R/get.file.R")
skel <- get.file("skel")
##
skume <- get.file("skume")
##
foli <- get.file("foli")
skel$urban <- FALSE
vb0Tab <- rbind(skel, skume[, names(skel)], foli[, names(skel)])
vb0Tab$civst <- relevel(vb0Tab$civst, ref = "married")
vb0Tab <- aggregate(vb0Tab[, c("event", "exposure")],
                    vb0Tab[, c("sex", "age", "period", "cohort", "civst", "hisclass", "urban")],
                    FUN = sum)
```

@@

```{r getske4060}
cap1w <-  "Cumulative hazards for social class by time period, ages 40--64, women."
scap1w <-  "Cumulative hazards, ages 40--64, women."
cap1m <-  "Cumulative hazards for social class by time period, ages 40--64, men."
scap1m <-  "Cumulative hazards, ages 40--64, men."
##

cap0 <- "Probability of dying before age 65 for a 40-year-old by social class and period, women (top) and men (bottom)."
scap0 <- "Probability of dying before age 65 for a 40-year old by social class."
cap1 <- "Probability of dying before age 90 for a 65-year-old by social class and period, women (top) and men (bottom)."
scap1 <- "Probability of dying before age 90 for a 65-year old by social class."
cap12 <- "Relative probability of dying before age 65 for a 40-year-old by social class and period, women (top) and men (bottom)."
scap12 <- "Relative probability of dying before age 65 for a 40-year-old by social class."
cap12log <- "Probability of dying before age 65 for a 40-year-old by social class and period, women (top) and men (bottom). The last decades."
scap12log <- "Probability by social class, ages 40--64, last years."

@@

```{r hazcalw}
source("R/getPHglm2.R")
source("R/getPHglm.R")
##nper <- 4
nper <- length(labb)
resw <- getPHglm(vb4064[vb4064$sex == "female", ],
            nper, labb)
resm <- getPHglm(vb4064[vb4064$sex == "male", ],
            nper, labb)
```

@@

```{r cumswstand, fig.cap = cap12, fig.scap = scap12, fig.height = 8, out.height = "14cm"}
oldpar <- par(mfrow = c(2, 1))
woutp <- plotED.glm(resw, col = scol, logScale = FALSE, relative = TRUE,
                    ylab = "P(dead at 65 | alive at 40)", ylim = c(0.0, 0.6), main = "Women 40-64")##, exclude = 5:6)
moutp <- plotED.glm(resm, col = scol, logScale = FALSE, relative = TRUE,
                    ylab = "P(dead at 65 | alive at 40)", ylim = c(0.0, 0.6), main = "Men 40-64")##, exclude = 5:6)
par(oldpar)
```

@@

```{r cumswold, fig.cap = cap1, fig.scap = scap1, fig.height = 8, out.height = "14cm"}
##source("R/plotED.R")
source("R/getPHglm.R")
resw <- getPHglm(vb6589[vb6589$sex == "female", ],
            nper, labb)
resm <- getPHglm(vb6589[vb6589$sex == "male", ],
            nper, labb)

oldpar <- par(mfrow = c(2, 1))
woutp <- plotED.glm(resw, col = scol, logScale = FALSE, relative = FALSE,
                    ylab = "P(dead at 90 | alive at 65)", ylim = c(0.4, 1.0), main = "Women 65-89")##, exclude = 5:6)
moutp <- plotED.glm(resm, col = scol, logScale = FALSE, relative = FALSE,
                    ylab = "P(dead at 90 | alive at 65)", ylim = c(0.4, 1.0), main = "Men 65-89")##, exclude = 5:6)
par(oldpar)
```

```{r cumsw2, fig.cap = cap22, fig.scap = scap22, fig.height = 8, out.height = "14cm"}
source("R/plotED.glm.R")
oldpar <- par(mfrow = c(2, 1))
woutp <- plotED.glm(resw, col = scol, logScale = FALSE, relative = TRUE, 
                    ylab = "P(dead at 90 | alive at 65)",
                main = "Women 65-89", ylim = c(0, 0.6))##, exclude = 5:6)
moutp <- plotED.glm(resm, col = scol, logScale = FALSE, relative = TRUE,
                    ylab = "P(dead at 90 | alive at 65)",
                main = "Men 65-89", ylim = c(0, 0.6))##, exclude = 5:6)
par(oldpar)
```
